---
title             : "Consistency and variability in word learning across languages"
shorttitle        : "Word learning consistency and variability"

author:
  - name          : Anonymized

affiliation:
  - institution  : ""

abstract: |
  Why do children learn some words earlier than others? The order in which words are acquired can provide clues about the mechanisms of word learning. In a large-scale corpus analysis, we use data from over 38,000 children to estimate the acquisition trajectories of around 400 words in ten languages, predicting them on the basis of independently-derived environmental and conceptual factors. We examine the consistency and variability of these predictors across languages, by lexical category, and over development. The ordering of predictors across languages is quite similar, suggesting similar processes in operation. In contrast, the ordering of predictors across different lexical categories is distinct, in line with theories that posit distinct factors at play in the acquisition of content words and function words. By leveraging data at a significantly larger scale than previous work, our analyses identify candidate generalizations about the processes underlying word learning across languages.

keywords          : word learning, language acquisition, corpus analysis

bibliography      : [aoapred.bib]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
csl               : pnas.csl
---

```{r include=FALSE} 
library(tidyverse)
library(magrittr)
library(langcog)
library(wordbankr)
library(forcats)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache = TRUE, dev = 'cairo_pdf')
# theme_set(theme_bw() +
#             theme(panel.grid = element_blank(),
#                   strip.background = element_blank(),
#                   legend.key = element_blank()))
theme_set(papaja::theme_apa())

label_caps <- function(value) {
  if_else(toupper(value) == value, value,
          paste0(toupper(substr(value, 1, 1)),
                 tolower(substr(value, 2, nchar(value)))))
}

data_path <- "../../aoa_unified/saved_data"
uni_joined <- feather::read_feather(
  file.path(data_path, "uni_joined.feather")
)
uni_model_data <- feather::read_feather(
  file.path(data_path, "uni_model_data.feather")
)
lang_coefs <- feather::read_feather(
  file.path(data_path, "lang_coefs.feather")
)
lexcat_coefs <- feather::read_feather(
  file.path(data_path, "lexcat_coefs.feather")
)

baseline_summary <- feather::read_feather(
  file.path(data_path, "baseline_summary.feather")
)
empirical_summary <- feather::read_feather(
  file.path(data_path, "empirical_summary.feather")
)
empirical_lang_summary <- feather::read_feather(
  file.path(data_path, "empirical_lang_summary.feather")
)

predictors <- levels(lang_coefs$term)
display_predictors <- predictors %>%
  map_chr(~gsub("_", " ", .x) %>% label_caps()) %>%
  set_names(predictors)

num_langs <- n_distinct(lang_coefs$language)
```

# Introduction

Despite tremendous individual variation in rate of development
[@fenson2007], the first words that children utter are strikingly consistent
[@tardif2008;@schneider2015]: they tend to talk about important people in their
life ("mom", "dad"), social routines ("hi", "uh oh"), animals ("dog", "duck"),
and foods ("milk", "banana"). Even as children learn from their experiences
and according to their own interests [@mayor2014;@nelson1973], their vocabulary
grows rapidly, typically adding more nouns, but also verbs ("go") and other
predicates ("hot") to their repertoires.  Over just their first three
years, children learn hundreds or even thousands of words
[@fenson1994;@mayor2011]. Why are some words learned so early and some much
later?

This simple question about the order of the acquisition of first words can
provide a window into the nature of children's language learning. Posed as a
statistical problem, the challenge is to find what variables best predict the
age at which different words are acquired. Previous work using this approach has
revealed that in English, within a lexical category (e.g., nouns, verbs), words
that are more frequent in speech to children are likely to be learned earlier
[@goodman2008]. And further studies have found evidence that a variety of other
semantic and linguistic factors are related to word acquisition
[@hills2009;@stokes2010;@perry2015;@roy2015;@swingley2017].

But these exciting findings are limited in their generality because they use
different datasets, focus on different predictors, and almost exclusively
analyze data from English-elearning children. They do not allow for cross-linguistic comparison of the relative importance of the many relevant factors under consideration in children from different language communities. Such cross-linguistic comparisons are
critical. Identifying commonalities (and differences) across languages is our
best strategy for uncovering the universal mechanisms that are in play for all
children and differentiating them from patterns of acquisition that emerge due
to the particulars of a given language or culture [@slobin1985;@bates1987]. Our
goal is thus to extend these classic approaches by assessing the degree to which the
predictors of word learning are consistent across different languages and
cultures, as well as whether there are similar patterns across different word
types (e.g., nouns vs. verbs).

The primary tool for characterizing the breadth of children's early vocabulary in these previous studies has been structured parent report. Naturalistic language samples and experimental methods are both valuable methods for assessing aspects of child language [@bornstein1998,@fernald2006]. But outside of a few ultra-dense transcripts [e.g., @roy2015], neither method typically provides the kind of holistic view that comes from parent report. We focus in particular on the MacArthur-Bates Communicative Development Inventory [CDI; @fenson2007], a family of parent-report vocabulary checklists in which parents are asked whether their child either "understands" or "says and understands" individual words.

CDIs are an inexpensive and widely-used method for gathering reliable
and valid data about the nature and extent of young children's productive and
receptive language [see @fenson1994 for review; cf. @feldman2000, @fenson2000]. Although CDIs cannot reveal the whole of a child's vocabulary, they do give an estimate of a child's knowledge about several hundred words, far more than the handful that are typically tested in a lab experiment. In addition, because of the low cost of adminstering survey instruments of this type compared with transcribing naturalistic recordings, it is far easier to gather samples in containing data about hundreds or thousands of children. Such large samples in turn make it possible to recover stable estimates of the average difficulty of individual words, even if individual children's data may be noisy.

Finally, CDI instruments have been used in dozens of different languages, providing an opportunity for cross-linguistic comparison. The American English CDI is not simply translated to other languages verbatim, however; instead, groups typically adapt the form for their particular linguistic and cultural situation as well as their scientific goal. This process leads to a wide range of forms with a common structure, but htat contain sets of words that are customized to particular languages and cultures to provide relevant information regarding how individual children's vocabularies vary (see SI for more details on the adaptations used in our study).

In our study, we conduct cross-linguistic comparisons of the age of acquisition of particular words using Wordbank
([wordbank.stanford.edu](http://wordbank.stanford.edu)) [@frank2016], an open repository that aggregates administrations of the CDI across languages. Wordbank currently includes data from more than 60,000 children. This large sample means that idiosyncratic patterns of individual variation are less likely to overpower general trends in which words are relatively easier or harder to learn, allowing us to investigate what features affect their acquisition across languages.


In our analyses, we integrate estimates of words' acquisition trajectories from the Wordbank data with independent characterizations of the word learning environment from other datasets. The use of secondary datasets for these analyses is warranted because no currently available resource provides data on *both* children's language environment and their learning outcomes for more than a small handful of children). 
We derive our estimates of the language environment from transcripts of speech to children in the CHILDES database
[@macwhinney2000]. This data-integration methodology wsa originated by @goodman2008; it relies on large samples to average out the (substantial) differences
between children and care environments, but it also allows for analyses that cannot be
performed on smaller datasets or datasets that measure only child or environment
but not both.

As our particular measures of environmental input, we estimated each word's
frequency, mean length of utterance in words (MLU), frequency as a sole
utterance constituent, and frequency in utterance-final position. While these
measures are crude, they are both easy to compute and relatively comparable
across the languages in our sample. We combined these measures with available
psycholinguistic norms for the words in our sample, using adult ratings of each
word's concreteness, valence, arousal, and association with babies as proxies
for conceptual properties of each word. We used these measures to predict words’
acquisition trajectories, assessing the relative contributions of each, as well
as how they change over development and interact with lexical category. 

These analyses address two  questions. First, we ask about the degree of consistency across languages in the relative importance of predictors. Consistency in the ordering of predictors would
suggest that similar information sources are important for learners, regardless
of language. Such evidence would point to the importance of similar learning
mechanisms across languages despite superficial dissimilarities (e.g., greater
morphological complexity in Russian and Turkish, greater phonological complexity
in Danish). Conversely, variability would reveal the degree to which learners
face different challenges in learning different languages.

Second, we ask which lexical categories are most influenced by environmental
factors, like frequency and utterance length, compared with conceptual factors
like concreteness and valence. Division of dominance theory suggests that nouns
might be more sensitive to conceptual factors, while predicates and closed-class
words might be more sensitive to environmental (linguistic) factors
[@gentner2001]. And on syntactic bootstrapping theories [@gleitman1990], nouns
are argued to be learned via frequent co-occurrence (operationalized by
frequency) while verbs might be more sensitive to syntactic factors
(operationalized here by utterance length) [@snedeker2007]. Thus, examining the
relative contribution of different predictors across lexical categories can help
test the predictions of influential theories of acquisition.


# Approach {#approach .unnumbered}

```{r lang_stats, results="asis"}
sample_sizes <- feather::read_feather(
  file.path(data_path, "sample_sizes.feather")
)

uni_lemmas <- uni_model_data %>%
  group_by(language) %>%
  summarise(num_included = n_distinct(uni_lemma))

lang_stats <- sample_sizes %>%
  spread(measure, num_admins) %>%
  left_join(uni_lemmas) %>%
  mutate(language = gsub("(.*) \\(.*\\)", "\\1", language)) %>%
  select(Language = language,
         "CDI Items" = num_included,
         "N production" = produces,
         "N comprehension" = understands)

lang_table <- xtable::xtable(
  lang_stats, digits = 0,
  caption = "Dataset statistics for acquisition data from Wordbank.",
  label = "table:lang_stats"
)

print(lang_table, type = "latex", comment = FALSE, include.rownames = FALSE,
      table.placement = "b",
      format.args = list(big.mark = ","))
```

To estimate the trajectory of words' acquisition, we used vocabulary data collected using CDI instruments adapted in many different languages, including both Words & Gestures (WG) forms for younger children and Words & Sentences (WS) forms for older children. When filling out a CDI form, parents are either asked to indicate whether their child "understands" (comprehension) or "understands and says" (production) each of around 400-700 words. Typically, both comprehension and production are queried for younger children and only production is queried for older children, but details vary from adaptation to adaptation. We use data from the items on the WG form for our comprehension measure, and data from the items in common between the WG and WS forms for our production measure. Table 1 gives an overview of our acquisition data [@kovacevic1996;@bleses2008;@boudreault2007;@trudeau2011;@caselli2012;@caselli1995;@simonsen2014;@vershinina2011;@yeliseyeva2009;@fenson2003;@eriksson2002;@acarlar2008].


```{r, fig.width=7, fig.height=3.5, fig.env='figure*', fig.align='center', fig.cap='Example production trajectories for the words "dog" and "jump" across languages. Points show average proportion of children producing each word for each one-month age group. Lines show the best-fitting logistic curve. Labels show the forms of the word in each language.'}
demo_lemmas <- c("dog", "jump")
demo_data <- uni_joined %>%
  filter(uni_lemma %in% demo_lemmas,
         measure == "produces") %>%
  mutate(language = gsub("(.*) \\(.*\\)", "\\1", language))

word_data <- demo_data %>%
  distinct(language, uni_lemma, words) %>%
  mutate(x = ifelse(uni_lemma == demo_lemmas[1], 8, 35),
         y = ifelse(uni_lemma == demo_lemmas[1], 1, 0))

ggplot(demo_data, aes(x = age, y = prop, colour = uni_lemma)) +
  facet_wrap(~language, ncol = 5) +
  geom_point(size = 0.8, alpha = 0.4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),
              se = FALSE, size = 1) +
  geom_label(aes(x = x, y = y, label = words), data = word_data, size = 3,
             label.padding = unit(0.15, "lines"),
             vjust = "inward", hjust = "inward") +
  scale_colour_solarized(guide = FALSE) +
  scale_y_continuous(name = "Proportion of children producing") +
  scale_x_continuous(name = "Age (months)", breaks = seq(10, 30, 10))
```

For each word, CDI data yield a trajectory reflecting the number of children that are reported to produce or understand the word at each age covered by the instrument (see Figure 1 for some examples). We then use a mixed-effects logistic regression model to predict whether each child knows the word on the basis of the child's age, properties of the word, and interactions between age and each property of the word. We also fit such models separately to the words in each lexical category. The magnitude of the standardized coefficient on each feature gives an estimate of its importance in predicting whether words are learned earlier or later. Interactions between features and age give estimates of how this effect is modulated for earlier and later-learned words. For example, a positive effect of association with babies ("babiness") means that words associated with babies are learned earlier; a negative interaction with age means that high babiness primarily leads to higher rates of production and comprehension for younger children.


```{r extremes, results='asis'}
num_extremes <- 3
extremes <- uni_model_data %>%
  filter(measure == "understands") %>%
  distinct_(.dots = c("language", "uni_lemma", predictors)) %>%
  mutate(uni_lemma = gsub("(.*) \\(.*\\)", "\\1", uni_lemma)) %>%
  split(.$language) %>%
  map_df(function(lang_data) {
    map_df(predictors, function(predictor) {
      if (predictor %in%
          c("frequency", "final_frequency", "solo_frequency", "MLU")) {
        filtered_lang_data <- lang_data %>%
          filter(frequency != min(frequency))
      } else {
        filtered_lang_data <- lang_data
      }
      highest <- filtered_lang_data %>%
        arrange_(as.formula(sprintf("~desc(%s)", predictor))) %>%
        .$uni_lemma %>%
        .[1:num_extremes]
      lowest <- filtered_lang_data %>%
        arrange_(predictor) %>%
        .$uni_lemma %>%
        .[1:num_extremes]
      return(data.frame("language" = unique(lang_data$language),
                        "Predictor" = predictor,
                        "highest" = paste(highest, collapse = ", "),
                        "lowest" = paste(lowest, collapse = ", ")))
    })
  })

extremes_display <- extremes %>%
  filter(language == "English") %>%
  select(-language) %>%
  rename(Lowest = lowest, Highest = highest) %>%
  mutate(Predictor = gsub("_", " ", Predictor),
         Predictor = map_chr(Predictor, label_caps)) %>%
  arrange(desc(row_number()))

extremes_table <- extremes_display %>%
  xtable::xtable(
    label = "table:extremes",
    caption = "Items with the highest and lowest values for each predictor in English."
  )

print(extremes_table, type = "latex", comment = FALSE,
      table.placement = "t",
      include.rownames = FALSE)
```


```{r, fig.width=7, fig.height=4.5, fig.env='figure*', fig.align='center', fig.cap="Estimates of coefficients in predicting words' developmental trajectories. Each point represents a predictor's coefficient in one language, with the large point showing the mean across languages. Larger coefficient values indicate a greater effect of the predictor on acquisition: positive main effects indicate that words with higher values of the predictor tend to be understood/produced by more children, while negative main effects indicate that words with lower values of the predictor tend to be understood/produced by more children; positive age interactions indicate that the predictor's effect increases with age, while negative age interactions indicate the predictor's effect decreases with age."}

num_coefs <- length(unique(lang_coefs$term))

lang_coefs <- lang_coefs %>%
  mutate(term = `levels<-`(term, display_predictors))

mean_coefs <- lang_coefs %>%
  group_by(term, measure, interaction) %>%
  summarise(mean = mean(estimate),
            ci_lower = ci_lower(estimate),
            ci_upper = ci_upper(estimate))

ggplot(lang_coefs, aes(x = estimate, y = term)) +
  facet_grid(measure ~ interaction, labeller = as_labeller(label_caps)) +
  geom_point(aes(colour = term), size = 1, alpha = 0.4) +
  geom_point(aes(x = mean, colour = term), size = 3, data = mean_coefs) +
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed") +
  scale_fill_solarized(guide = FALSE) +
  scale_colour_manual(guide = FALSE,
                      values = rev(solarized_palette(num_coefs))) +
  ylab("") +
  scale_x_continuous(name = "Coefficient estimate")
```

Each of our predictors is derived from independent sources. For each word that appears on the CDI forms in each of our `r num_langs` languages, we used corpora of child-directed speech in that language to obtain an estimate of its frequency, the mean length of utterances in which it appears, its frequency as the sole consituent of utterance, and its frequency in utterance final position (with frequency residualized out of solo and final frequencies). Additionally, we computed each word's length in characters and included ratings of its concreteness, valence, arousal, and relatedness to babies. 

We included this last factor because previous studies have shown robust consistency in the types of words that children learn very early [@tardif2008]. These words seem to describe concepts that are important or exciting in the lives of infants in a way that standard psycholinguistic features like concreteness do not. Capturing this intution quantitatively is difficult, but @perry2015 provides a proxy measure as a first step. This measure is simply the degree to which a particular word was "associated with babies." Intuitively, we expect this measure to capture the degree to which words like "bottle" feature heavily in the environment (and presumably, mental life) of many babies. 

Since existing datasets for all of these conceptual ratings are primarily available for English, we mapped all words onto translation equivalents across CDI forms, verified by native speaker judgements, allowing us to use the ratings for English words across languages. While necessarily imperfect, this method allows us to examine languages for which limited resources exist. Example words for these predictors in English are shown in Table 2.

```{r collinearity}
cors <- uni_model_data %>%
  split(paste(.$language, .$measure)) %>%
  map_df(function(lang_data) {
    pred_data <- lang_data %>%
      select_(.dots = c("uni_lemma", predictors)) %>%
      distinct() %>%
      gather_("predictor", "value", predictors) %>%
      split(.$predictor)
    
    map(pred_data, function(pd1) {
      map_dbl(pred_data, function(pd2) cor(pd1$value, pd2$value))
    }) %>%
      as.data.frame() %>%
      mutate(predictor1 = row.names(.)) %>%
      gather_("predictor2", "cor", predictors) %>%
      filter(predictor1 != predictor2) %>%
      mutate(language = unique(lang_data$language),
             measure = unique(lang_data$measure))

  })

mean_cors <- cors %>%
  group_by(predictor1, predictor2) %>%
  summarise(mean_cor = mean(cor))
pair_cor <- function(p1, p2) {
  mean_cors %>%
    filter(predictor1 == p1, predictor2 == p2) %$%
    mean_cor
}
```

A potential concern for comparing coefficient estimates is predictor collinearity. Fortunately, in every language, the highest correlations were between MLU and solo frequency (mean over languages and measures _r = _ `r sprintf("%.2f", pair_cor("solo_frequency", "MLU"))`), as expected given the similarity of these factors; and frequency and number of characters (mean over languages and measures _r = _ `r sprintf("%.2f", pair_cor("frequency", "length"))`), a reflection of Zipf's Law [@zipf1935].


# Results {#results .unnumbered}

Figure 2 shows the coefficient estimate for each predictor in each language. We find that babiness, frequency, MLU, and concreteness are relatively stronger predictors of age of acquisition across languages. Given the emphasis on frequency effects in the language acquisition literature [@ambridge2015], one might have expected frequency to dominate, but several other predictors are just as strong in this analysis. Some factors previously argued to be important for word learning, namely valence and arousal [@moors2013], appear to have limited relevance when compared to other factors.


```{r, fig.width=5, fig.height=3, fig.align='center', fig.cap="Correlations of coefficients estimates between languages. Each point represents the mean of one language's coefficients' correlation with each other language's coefficients, with the black line indicating the overall mean across languages. The grey region and dashed line show a bootstrapped 95\\% confidence interval of a randomized baseline where predictor coefficients are shuffled within language."}
ggplot(NULL, aes(x = dummy, y = mean_cor)) +
  facet_grid(measure ~ interaction, labeller = as_labeller(label_caps)) +
  coord_flip() +
  geom_point(aes(y = cor, colour = language),
             data = empirical_lang_summary, alpha = 0.5) +
  geom_errorbarh(aes(xmin = 0.8, xmax = 1.2, x = 1, y = cor),
                 size = 1, height = 0,
                 data = empirical_summary) +
  geom_rect(aes(ymin = ci_lower_cor, ymax = ci_upper_cor, xmin = 0.5,
                xmax = 1.5), data = baseline_summary, colour = "lightgray",
            alpha = .1, linetype = 0) +
  geom_hline(data = baseline_summary, aes(yintercept = mean_cor), lty = 2) + 
  scale_y_continuous(name = "Mean pairwise correlation of coefficients",
                     limits = c(-0.1, 1)) +
  xlab("") +
  scale_colour_solarized(guide = FALSE) +
  theme(axis.ticks.y = element_blank()
        #axis.title.y = element_text(size = 14),
        #strip.text.y = element_text(size = 6)
        )
```


```{r, fig.width=7, fig.height=3.5, fig.env='figure*', fig.align='center', fig.cap="Estimates of coefficients in predicting words' developmental trajectories (as described in Figure 2), with separate models for each lexical category."}
plt_lexcat_coefs <- lexcat_coefs %>%
  filter(interaction == "main effect",
         !(term %in% c("valence", "arousal"))) %>%
  mutate(term = `levels<-`(term, display_predictors))

mean_lexcat_coefs <- plt_lexcat_coefs %>%
  group_by(lexical_category, term, measure, interaction) %>%
  summarise(mean = mean(estimate),
            ci_lower = ci_lower(estimate),
            ci_upper = ci_upper(estimate))

ggplot(plt_lexcat_coefs, aes(x = term, y = estimate)) +
  facet_grid(measure ~ lexical_category, labeller = as_labeller(label_caps)) +
  geom_point(aes(colour = term), size = 1, alpha = 0.4) +
  geom_point(aes(y = mean, colour = term), size = 3,
             data = mean_lexcat_coefs) +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
  coord_flip() +
  scale_fill_solarized(guide = FALSE) +
  scale_colour_manual(guide = FALSE,
                      values = rev(solarized_palette(num_coefs)),
                      drop = FALSE) +
  labs(x = "", y = "Coefficient estimate")
```


Overall, there is considerable consistency in the magnitudes of predictors across languages. In almost all, babiness and frequency were highest, while valence and arousal were smaller. A priori it could have been the case that different languages have wildly different effects of experiential vs. structural factors, but this pattern is not what we observe. Instead, Figure 3 shows the mean pairwise correlation of predictor coefficients across languages (i.e., the correlation of coefficients for English with coefficients for Russian, for Spanish, and so on). These means -- and even the individual datapoints -- are far outside of bootstrapped estimates for the average pairwise correlation in a randomized baseline created by shuffling predictor coefficients within language, suggesting that coefficient estimates are far more consistent across languages than would be expected by chance.

Word length is the one predictor of acquisition that varied substantially between measures, in that it is far more predictive for production than comprehension. Thus as measured here, length seems to be playing more the role of production constraints (i.e., how difficult a word is to say) than comprehension constraints (i.e., how difficult it is to store or access).

Next, we wanted to examine how the relative contributions of the predictors changes over development. Across languages, positive age interactions can be seen for concreteness and frequency (i.e., their effects increase with age). Conversely, there are negative age interactions for babiness and valence in comprehension and for solo frequency in production, suggesting stronger effects in words learned earlier in development.

Previous work gives reason to believe that predictors' relationship with age of acquisition differs among various lexical categories [@goodman2008]. To investigate these effects, we separated our data by lexical category and fit separate models for each category. Figure 4 shows the resulting coefficient estimates. Across languages, frequency had the highest magnitude for nouns and a lower magnitude for function words. In contrast, MLU was almost irrelevant for both nouns and predicates, but highly predictive for function words. These patterns are supportive of the hypothesis that different word classes are learned in different ways, or at least that the bottleneck on learning tends to be different, leading to different information sources being more or less important across categories.

# Discussion {.unnumbered}

What makes words easier or harder for young children to learn? Previous
experimental work has largely addressed this question using small-scale
experiments. While such experiments can identify sources of variation, they
typically do not allow for different sources to be compared in detail. In
contrast, observational studies allow the effects of individual factors to be
measured across ages and lexical categories [e.g.,
@goodman2008;@hills2009;@swingley2017]. Such work has identified a number of
candidate predictors of word learning. By including `r num_langs` languages and
`r num_coefs` predictors, our work expands the scope of these studies
dramatically, leading to several new findings.

First, we found consistency in the ordering of predictors across languages at a
level substantially greater than the predictions of a chance model. This
consistency supports the idea that differences in culture or language structure
do not lead to fundamentally different acquisition strategies, at least at the
level of detail we were able to examine. Instead, they are likely produced by
processes that are similar across populations and languages. Such processes
could include learning mechanisms or biases internal to children, or
interactional dynamics between children or caregivers. We believe these
consistencies should be an important topic for future investigation.

Second, predictors varied substantially in their weights across lexical
categories. Frequent, concrete nouns were learned earlier, consistent with
theories that emphasize the importance of early referential speech [e.g.,
@baldwin1995]. But for predicates, concreteness was somewhat less important. And
for function words, MLU was most predictive, perhaps because it is easiest to
decode the meanings of function words that are used in short sentences (or
because such words have meanings that are easiest to decode). Overall these
findings are consistent with some predictions of both division of dominance
theory, which highlights the role of conceptual structure in noun acquisition
[@gentner2001], and syntactic bootstrapping theory, which emphasizes linguistic
structure over conceptual complexity in the acquisition of lexical categories
other than nouns [@snedeker2007]. More generally, our methods here provide a way
forward for testing the predictions of these theories across languages and at
the level of the entire lexicon rather than individual words.

In addition to these new insights, several findings emerge that confirm and
expand previous reports. Environmental frequency was an important predictor of
learning, with more frequently heard words learned earlier
[@goodman2008;@swingley2017]. Predictors also changed in relative importance
across development. For example, certain words whose meanings were more strongly
associated with babies appeared to be learned early for children across the
languages in our sample [as in @tardif2008]. Finally, word length showed a
disassociation between comprehension and production, suggesting that challenges
in production do not carry over to comprehension (at least in parent-report
data).

Despite its larger scope, our work shares a number of important limitations with
previous studies. First and foremost, our approach is to predict one set of
individuals with data about the experience of a completely different set and
ratings of concepts gathered from yet others. In contrast to dense-data analyses
[@roy2015], this approach fundamentally limits the amount of variability we will
be able to capture. In addition, the granularity of the predictors that can be
extracted from corpus data and applied to every word is necessarily quite
coarse. Ideally, predictors could be targeted more specifically at particular
theoretical constructs of interest (for example, the patterns of use for
specific predicates).

Finally, our data are observations gleaned from parent report. CDI instruments are both reliable and valid, and the cross-linguistic adaptations we used contain the original researchers' best attempts to create culturally-appropriate word lists. Nevertheless,  several biases are still likely present in our data. First, parent 


Thus, our conclusions will require further testing through converging evidence
from both laboratory experiments and direct observation.

<!-- Finally, our work underscores the incompleteness of the current
understanding of vocabulary development. Even for English, the language in which
our model captures the most variance ($r^2 = 0.29$), much still remains
unexplained. Furthermore, this variance is highly reliable---cross-validation
using half of the English-speaking children to predict ages of acquisition for
the other half yields $r^2 = 0.98$. This gap highlights an important theoretical
challenge in the study of early language: linking individual datapoints to the
broader patterns of acquisition. We have strong theories of how individual
learning situations proceed, but must unify these theories to make progress on
understanding language learning at scale. -->

In sum, by examining predictors of early word learning across languages, we
identified substantial cross-linguistic consistency in the factors contributing
to the ease or difficulty of learning individual words. These findings testify
to the importance of building open, shared resources in the study of language
learning -- without the efforts of many research groups across many language
communities, such studies would be impossible. In addition, we hope that our
work here provides a baseline for the building of future predictive models that
allow theories of language learning to be tested at scale.



```{r}
missing <- uni_joined %>%
  select_(.dots = c("language", "uni_lemma", predictors)) %>%
  distinct() %>%
  gather_("predictor", "value", predictors) %>%
  filter(is.na(value)) %>%
  group_by(language, predictor) %>%
  summarise(n = n())
```

\newpage

# Materials and Methods

All code and data to reproduce our analyses are available at `https://github.com/mikabr/aoa-prediction`.

## Predictor variables

Each numeric predictor was centered and scaled so that all predictors would have comparable units. For each predictor, missing values (CDI items that were not in the relevant corpus or norms) were imputed from the mean for their respective language and measure. Placeholder items, such as "child's own name", were excluded.

Translation equivalents are available in the Wordbank database using the `wordbankr` package in `R` [@frank2016]. Translation equivalents were constructed by the authors and independently hand-checked by native speakers.

### Frequency

For each language, we estimated word frequency from unigram counts based on all corpora in CHILDES for that language. Each word's count includes the counts of words that share the same stem (so that "dogs" counts as "dog") or are synonymous (so that "father" counts as "daddy"). For polysemous word pairs (e.g., "orange" as in color or fruit), occurrences of the word in the corpus were split uniformly between the senses on the CDI. Counts were normalized to the length of each corpus, Laplace smoothed, and then log transformed.

### Solo and Final Frequencies

Using the same dataset as for frequency, we estimated the frequency with which each of word occurs as the sole word in an utterance, and the frequency with which it appears as the final word of an utterance (not counting single-word utterances). As with frequency, solo and final counts were normalized to the length of each corpus, Laplace smoothed, and log transformed. Since both of these estimates are by necessity highly correlated with frequency, we then residualized unigram frequency out of both of them, so that values reflect an estimate of the effect of solo/final frequency over and above frequency itself.

### MLU

For each language, we estimated each word's MLU by calculating the mean length in words of the utterances in which that word appeared, for all corpora in CHILDES for that language. For words that occurred fewer than 10 times, MLU estimates were not used (i.e. treated as missing).

### Length

We computed the number of characters in each word in each language. While imperfect, this metric of length is highly correlated with number of phonemes and syllables [@lewis2016].

### Concreteness

We used previously collected norms for concreteness [@brysbaert2014], which were gathered by asking adult participants to rate how concrete the meaning of each word is on a 5-point scale from abstract to concrete.

### Valence and Arousal

We also used previously collected norms for valence and arousal [@warriner2013], for which adult participants were asked to rate words on a 1-9 happy-unhappy scale (valence) and 1-9 excited-calm scale (arousal).

### Babiness

Lastly, we used previously collected norms of "babiness", a measure of association with infancy [@perry2015] for which adult participants were asked to judge a word's association with babies on a 1-10 scale.


## Lexical category

Category was determined on the basis of the conceptual categories presented on the CDI form (e.g., "Animals"), such that the Nouns category contains common nouns, Predicates contains verbs and adjectives, Function Words contains closed-class words, and Other contains the remaining items [following @bates1994].


## Analysis

For all analyses, we used logistic mixed-effects regression models (fit with `lme4 1.1-12` in `R`) to predict whether each child understands/produces each word from age, the above predictors, and the interactions of age with the predictors. Each model was fit to all data from a particular language community and included a random intercept for each word and a random slope of age for each word.


# Acknowledgements

Thank you to the labs and individuals who contributed data to Wordbank and to NSF BCS \#1528526 for support.


\newpage

# References

```{r create_r-references}
papaja::r_refs(file = "aoapred.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
