---
title: |
  | From _uh-oh_ to _tomorrow_
  | Predicting age of acquisition for early words across languages
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Mika Braginsky} \\ \texttt{mikabr@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Daniel Yurovsky} \\ \texttt{yurovsky@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Virginia A. Marchman} \\ \texttt{marchman@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

abstract: 
    "_[TODO: abstract]_"
    
keywords:
    "language acquisition; word learning; development"
    
output: cogsci2016::cogsci_paper
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width = 3, fig.height = 3, fig.crop = F,
                      fig.pos = "tb", fig.path = 'figs/', echo = F,
                      warning = F, cache = F, message = F, sanitize = T)
```

```{r, libraries}
library(knitr)
library(png)
library(grid)
library(xtable)
library(ggplot2)
library(langcog)
library(dplyr)
library(tidyr)
library(purrr)
# library(extrafont)
# loadfonts(quiet = TRUE)
font <- "Times"
theme_set(theme_mikabr(base_size = 12, base_family = font) +
            theme(legend.key = element_blank(),
                  legend.background = element_rect(fill = "transparent"),
                  legend.margin = unit(0, "mm"),
                  panel.background = element_blank(),
                  panel.margin = unit(1, "mm"),
                  plot.margin = unit(c(0, 0, 0, 0), "mm")))
```

```{r}
# ----------
# to run and cache all the AoA prediction computation:
# - set working directory to this file's location
# - uncomment and run these four lines
# ----------
# setwd("..")
# knit("aoa_prediction_funs.Rmd", tangle = TRUE)
# source("aoa_prediction_funs.R")
# save.image("aoa_prediction.RData")

# when knitting, just uses cache
load("../aoa_prediction.RData")
```

# Introduction

Our approach to addressing these questions is to use large-scale vocabulary development data to examine the contribution of various theoretically-relevant factors to vocabulary growth. We can look across children to determine how easy or hard various words are to learn, and then examine the relation between word difficulty and various word properties that relate to proposed word-learning factors. Foundation work using such an approach has revealed that in English, within lexical category, words that are more frequent in speech to children are likely to be learned earlier [@goodman2008; @roy2009]. Further studies have delved into the relevance of semantic networks [@hills2009], neighborhood density [@stokes2010], iconicity [@perry2015], and linguistic distinctiveness [@roy2015predicting] as possible contributors to vocabulary development.

These previous studies used different datasets, focused on different predictors, and for the most part, only analyzed English data. It is thus impossible to compare the relative importance of the many relevant factors under consideration and to draw robust conclusions. To remedy this issue, we present analyses based on data from Wordbank (wordbank.stanford.edu), an open repository of cross-linguistic language development data [@frankinpress]. By aggregating administrations of the MacArthur-Bates Communicative Development Inventory [CDI; @fenson2007], a family of parent-report vocabulary checklists, Wordbank provides large-scale vocabulary data based on analogous instruments from more than 40,000 children in 14 different language communities.  As such, Wordbank offers a novel resource for richer and more powerful analyses of vocabulary learning over development and across languages.

Here, we integrate Wordbank data with characterizations of the word learning environment from the CHILDES database [@macwhinney2000] and elsewhere, a multiple data source approach pioneered by @goodman2008. Building on their work, we want to move beyond frequency to examine a variety of information sources. We specifically follow @roy2015 in predicting age of acquisition (AoA) as a function of several different environment predictors. In analyzing a high-density longitudinal corpus for a single English-acquiring child, Roy et al found that frequency, number of characters, and mean length of utterances predicted the age of a word's first production. Due to the nature of the data, this analysis was limited to one language (in fact to one subject) and could only test production, distancing the connection between properties of the input and the child's emerging understanding of words.

Our approach provides a complimentary analysis by using CDI comprehension data available in Wordbank to look at a common set of the earliest words that children learn across several different languages. We estimate the age of acquisition for around 400 words listed on the CDIs in each of `r length(unique(lang_coef$language))` languages. We also estimate each word's frequency and mean length of utterance (MLU) based on all sentences in which each word appears in naturalistic corpora of talk to children [IS THIS TRUE?] in CHILDES. We also obtain ratings of each word's concreteness, valence, arousal, and relevance to babies from previously collected norms [CITATIONS??]. We then predict each word's AoA from all of these properties, assessing the relative contributions of each factor, as well as the interaction of predictors with the lexical category and how those interactions change over development. Each of these analyses has the potential to provide leverage on long-standing theoretical questions. 

A first theoretical question of interest is which lexical categories are most influenced by input-related factors like frequency and utterance length compared with conceptual factors like concreteness or valence. For example, the "division of dominance" theory suggests that nouns might be more sensitive to cognitive factors while predicates and closed-class words might be more sensitive to linguistic factors [@gentner2001]. On the other hand, on syntactic bootstrapping theories, nouns are argued to be learned via frequent co-occurrence (operationalized by frequency) while verbs might be more sensitive to syntactic factors (operationalized here by utterance length) [@gleitman1990], and neither would be particularly sensitive to conceptual complexity [@snedeker2007]. 

A second question of interest is the extent to which there is variability across languages in the relative importance of predictors. For example, are there differences in the importance of syntactic factors in morphologically more complex languages like Russian and Turkish, compared with simpler ones like English? Differences of this type might be revealing of the degree to which learners face different challenges in different language environments. Or consistency may suggest the operation of similar learning mechanisms and strategies that are not as dependent on the complexities of phonology, morphology, and syntax in a particular language. 

Overall, by incorporating a variety of theoretically-important factors, as well as basing our analysis on a large samples of words and children and building towards more cross-linguistic coverage, our study presents a more thorough investigation of the question of what properties determine words' learnability.

# Data

We use Wordbank (wordbank.stanford.edu), an open database of developmental vocabulary data, to estimate the age of acquisition for words across `r length(unique(lang_coef$language))` languages: `r paste(unique(lang_coef$language), collapse = ", ")`. We then ask what factors are most important for predicting this age of acquisition.

## Estimating Age of Acquisition

To estimate the age at which words are acquired, we took vocabulary data collected using the MacArthur-Bates Communicative Development Inventory, a family of parent-report checklists, specifically the Words & Gestures (infant) form for 8- to 18-month-olds. When filling out a CDI a form, parents are asked to indicate whether their child understands and/or says each of around 400 words. From these data, for each word on the CDI, we computed the proportion of children at each age that are reported to understand the word. We then fit a logistic curve to these proportions using a robust generalized linear model (using the \texttt{robustbase} package in \texttt{R}) and determine when the curve crosses 0.5, i.e. at what age at least 50% of children are reported to understand the word. Following @goodman2008, we take this point to be each word's age of acquisition.

```{r lang_stats, results="asis"}
library(wordbankr)
num_admins <- get_administration_data(mode = "remote") %>%
  filter(form == "WG") %>%
  group_by(language) %>%
  summarise(num_admins = n())

# num_words <- words %>%
#   group_by(language) %>%
#   summarise(num_words = n())

aoa_stats <- crossling_model_data %>%
  group_by(language) %>%
  summarise(num_included = n())
            #mean_aoa = mean(aoa),
            #var_aoa = var(aoa))

childes_stats <- read.csv("../predictors/frequency/num_words.csv") #%>%
  #mutate(num_words = formatC(num_words, big.mark = ","))

lang_stats <- aoa_stats %>%
  #left_join(num_words) %>%
  left_join(num_admins) %>%
  left_join(childes_stats)
colnames(lang_stats) <- c("Language", "CDI Items", "CDI Admins", "CHILDES Words")
                          #"Mean AoA",, "Var AoA")

lang_table <- xtable(lang_stats, digits = 2, caption = "Dataset statistics")

print(lang_table, type = "latex", comment = FALSE, include.rownames = FALSE,
      table.placement = "t", format.args = list(big.mark = ","))
```

```{r extremes, results="asis"}
num_extremes <- 3
extremes <- crossling_model_data %>%
    split(.$language) %>%
    map_df(function(lang_data) {
    map_df(c("aoa", rev(levels(crossling_predictors_ordered))), function(predictor) {
      highest <- lang_data %>%
        arrange_(as.formula(sprintf("~desc(%s)", predictor))) %>%
        .$uni_lemma %>%
        .[1:num_extremes]
      lowest <- lang_data %>%
        arrange_(predictor) %>%
        .$uni_lemma %>%
        .[1:num_extremes]
      return(data.frame("language" = unique(lang_data$language),
                        "Measure" = predictor,
                        "Lowest" = paste(lowest, collapse = ", "),
                        "Highest" = paste(highest, collapse = ", ")))
    })
  }) %>%
  filter(language == "English") %>%
  select(-language) %>%
  mutate(order = row_number()) %>%
  gather(Direction, Words, Lowest, Highest) %>%
  arrange(order) %>%
  select(-order, -Direction) %>%
  mutate(Measure = unlist(transpose(list(unique(Measure), rep("", length(Measure))))))

extremes_table <- xtable(extremes, label = "tab:mytable", #align = "llp{1in}p{1in}",
                         caption = "Examples of words with the lowest and highest values for age of acquisition and each predictor.")

print(extremes_table, type = "latex", comment = FALSE, include.rownames = FALSE,
      table.placement = "!hb", floating.environment = "table")
```

```{r data, fig.height=3.5, fig.width=7, fig.env='figure*', fig.align='center', fig.pos='!h', dev='cairo_pdf', fig.cap='Relationship between predictors and AoA for each lexical category in each language.'}
crossling_model_data %>%
  gather_("predictor", "value", crossling_predictors) %>%
  mutate(predictor = factor(predictor, levels = crossling_predictor_levels),
         lexical_category = factor(lexical_category,
                                   labels = paste(levels(lexical_category), "  "))) %>%
  filter(language %in% c("English", "Italian", "Norwegian")) %>%
  ggplot(aes(x = value, y = aoa)) +
    #facet_wrap(~predictor, ncol = 4) +
    facet_grid(language ~ predictor, scales = "free_x") +
    geom_point(aes(fill = lexical_category), colour = NA,
               size = 0.5, alpha = 0.4, shape = 21) +
    geom_smooth(aes(colour = lexical_category), method = "lm", se = FALSE, size = 0.8) +
    geom_smooth(color = "black", method = "lm", se = FALSE, size = 0.8) +
    scale_colour_solarized(name = "") +
    scale_fill_solarized(name = "") +
    xlab("Predictor z-Score") +
    scale_y_continuous(name = "Age of Acquisition (months)",
                       breaks = seq(5, 25, 5)) +
    #theme_mikabr(base_size = 10) +
    theme(legend.position = "top")
```

```{r}
predict_english <- function(english_predictors) {
  english_model_data <- english_model_data_fun(english_data_fun(uni_joined,
                                                                english_predictors,
                                                                num_characters),
                                               english_predictors)
  english_predictions_fun(english_model_data,
                          english_model_fun(english_model_data, english_predictors))
}
english_predictors <- c("frequency",  "concreteness", "num_characters",
                        "valence", "arousal", "babiness")
predictions <- predict_english(english_predictors)
predictions <- predictions %>%
    mutate(lexical_category = factor(lexical_category,
                                     levels = levels(lexical_category),
                                     labels = paste0(levels(lexical_category), "  ")),
           error = abs(predicted_aoa - aoa))

cors <- predictions %>%
  group_by(lexical_category) %>%
  summarise(cor = cor(aoa, predicted_aoa))
```

```{r fit, fig.height=2.3, fig.width=7, fig.env='figure*', fig.align='center', dev='cairo_pdf', fig.cap='English model fit.'}
threshold <- 6
predictions %>%
  ggplot(aes(x = predicted_aoa, y = aoa)) +
    facet_wrap(~lexical_category, ncol = 4) +
    geom_point(aes(colour = lexical_category), size = 0.7,
               data = filter(predictions, error < threshold)) +
    geom_label(aes(label = uni_lemma, colour = lexical_category), family = font,
              show.legend = FALSE, data = filter(predictions, error >= threshold),
              label.padding = unit(0.15, "lines"), size = 3) +
    geom_smooth(aes(colour = lexical_category), weight = 1, method = "lm", se = FALSE) +
    geom_smooth(colour = "black", weight = 2, method = "lm") +
    scale_colour_solarized(name = "", guide = FALSE) +
    scale_x_continuous(name = "Model Predicted AoA (months)",
                       breaks = seq(5, 25, by = 5), limits = c(5, 25)) +
    scale_y_continuous(name = "AoA (months)",
                       breaks = seq(5, 25, by = 5), limits = c(5, 25)) +
    geom_text(aes(x = 8, y = 24.5, label = sprintf("r = %.2f", cor)),
              data = cors, family = font, size = 2.8)
```

```{r devo, fig.height=3, fig.width=3, fig.env='figure', fig.align='center', fig.pos='!hb', dev='cairo_pdf', fig.cap='Predictor values over development.'}
crossling_data %>%
  gather_("predictor", "value", crossling_predictors) %>%
  filter(!is.na(value)) %>%
  mutate(predictor = factor(predictor, levels = crossling_predictor_levels,
                            labels = paste(crossling_predictor_levels, " "))) %>%
  ggplot(aes(x = aoa, y = value, colour = predictor)) +
    #facet_wrap(~predictor, ncol = 4) +
    #geom_point(size = 0.5, alpha = 0.2) +
    #geom_smooth(method = "loess", se = FALSE, span = 0.75) +
    geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
    geom_smooth(method = "lm", formula = y ~ poly(x, 3), se = FALSE) +
    scale_colour_solarized(name = "", guide = guide_legend(nrow = 4, byrow = FALSE)) +
    scale_x_continuous(name = "Age of Acquisition (months)",
                       breaks = seq(0, 30, 5)) +
    scale_y_continuous(name = "Predictor Z-Score",
                       breaks = seq(-1.5, 1.5, 0.5)) +
    theme(legend.position = c(0.64, 0.13),
          legend.direction = "horizontal",
          legend.key.height = unit(0.7, "lines"),
          legend.key.width = unit(0.7, "lines"),
          legend.text = element_text(size = 8))
```

```{r coefs, fig.height=3.5, fig.width=7, fig.env='figure*', fig.align='center', dev='cairo_pdf', fig.cap='Magnitudes of predictor coefficients.'}
plot_coefs <- crossling_coef %>%
  mutate(language = "Cross-Linguistic") %>%
  bind_rows(lang_coef) %>%
  mutate(language = factor(language, levels = c("Cross-Linguistic", langs)),
         term = factor(term, levels = rev(levels(term))),
         sign = factor(sign(estimate)))

ggplot(plot_coefs, aes(x = term, y = abs(estimate))) +
  facet_wrap(~language, ncol = 4) +
  geom_rect(data = filter(plot_coefs, language == "Cross-Linguistic"),
            aes(fill = language), xmin = -Inf, xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.03) +
  geom_linerange(aes(colour = term,
                     ymin = abs(estimate) - 1.96 * std.error,
                     ymax = abs(estimate) + 1.96 * std.error)) +
  geom_point(aes(colour = term, shape = sign), size = 2.5) +
#     geom_pointrange(aes(ymin = abs(estimate) - 1.96 * std.error,
#                         ymax = abs(estimate) + 1.96 * std.error,
#                         shape = sign)) +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
#    geom_text(aes(label = paste("bar(R)^2==", adj_rsq)), parse = TRUE,
  geom_text(aes(label = paste("bar(R)^2==", adj_rsq)), parse = TRUE,
            x = 1, y = 1.8, family = font, size = 2.8,
            data = data.frame(language = names(all_adj_rsq),
                              adj_rsq = round(unlist(all_adj_rsq), 2),
                              row.names = NULL)) +
  coord_flip() +
  scale_fill_solarized(guide = FALSE) +
  scale_colour_manual(guide = FALSE, values = rev(solarized_palette(7))) +
  scale_shape_manual(name = "", values = c(1, 19),
                     labels = c("negative", "positive")) +
  xlab("") +
  scale_y_continuous(name = "Coefficient Magnitude (Months/SD)",
                     breaks = seq(0, 2, 0.5)) +
  theme(legend.position = c(0.87, -0.1),
        legend.direction = "horizontal",
        legend.key.height = unit(0.5, "lines"),
        legend.key.width = unit(0.5, "lines"))
```

```{r coefs_lexcat, fig.height=1.7, fig.width=7, fig.env='figure*', fig.align='center', dev='cairo_pdf', fig.cap='Magnitudes of predictor coefficients by lexical category.'}
crossling_coefs %>%
  filter(term %in% c("frequency", "babiness", "concreteness", "mlu")) %>%
  mutate(term = factor(term, levels = rev(crossling_predictor_levels)),
         sign = factor(sign(estimate))) %>%
  ggplot(aes(x = term, y = abs(estimate), colour = term)) +
    facet_wrap(~lexical_category, ncol = 4) +
    geom_linerange(aes(ymin = abs(estimate) - 1.96 * std.error,
                        ymax = abs(estimate) + 1.96 * std.error)) +
    geom_point(aes(shape = sign), size = 2.5) +
#     geom_pointrange(aes(ymin = abs(estimate) - 1.96 * std.error,
#                        ymax = abs(estimate) + 1.96 * std.error)) +
    geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
    coord_flip() +
    scale_colour_manual(guide = FALSE, values = rev(solarized_palette(7))[4:7]) +
    scale_shape_manual(name = "", values = c(1, 19),
                       labels = c("negative", "positive")) +
    xlab("") +
    scale_y_continuous(name = "Coefficient Magnitude (Months/SD)",
                       breaks = seq(0, 2.5, 0.5)) +
    theme(legend.position = c(0.87, -0.26),
          legend.direction = "horizontal",
          legend.key.height = unit(0.5, "lines"),
          legend.key.width = unit(0.5, "lines"))
```

## Predictors

Each of our predictors are derived from independent resources. For each word that appears on the CDI Word & Gestures form in each of our `r length(unique(lang_coef$language))` languages, we obtained an estimate of its frequency in child-directed speech, the mean utterance length (MLU) of sentences in which if appears in child-directed speech, its length in characters, and ratings of its concreteness, valence, arousal, and relevance to babies. Examples of each of these predictors for English are shown in Table \ref{tab:extremes}.

While frequency and MLU are measured relative to the word's language, since the conceptual ratings were collected only in English, we mapped all the words onto translation equivalents across CDI forms, allowing us to use the ratings for English words cross-linguistically. While imperfect, this method allows us to examine languages for which limited resources exist. Translation equivalents are available in the wordbank database using the \texttt{wordbankr} package [@frankinpress].

Items such as _child's own name_ were excluded in all languages. Each predictor was also centered and scaled so that all predictors would have comparable units. Lexical category was determined on the basis of the conceptual categories presented on the CDI form (e.g., "Animals"), such that Nouns contains common nouns, Predicates contains verbs and adjectives, Function Words contains closed-class words, and Other contains the remaining items [following @bates1994].

### Frequency 

For each language, we estimated word frequency from unigram count in all corpora in the CHILDES database for that language, normalized to the length of the corpus. Each word's count includes the counts of words that share the same stem (so that _dogs_ counts as _dog_) or are synonymous (so that _father_ counts as _daddy_). For polysemous word pairs, such as _orange_ as in color and _orange_ as in fruit, each occurrence of _orange_ in the corpus counts for both. Finally, each word's frequency estimate is taken as the log of its count.

### MLU

For each language, we estimated each word's MLU by calculating the mean number of words in the sentences in which that word appears in all corpora in the CHILDES database for that language. Words that only occur in one sentence were excluded.

### Length

We computed the number of characters in each word in each language, which is known to be highly correlated with number of phonemes and syllables.

### Concreteness

We used previously collected norms for concreteness [@brysbaert2014], which were gathered by asking adult participants to rate how concrete the meaning of each word is by using a 5-point scale from abstract to concrete. For the `r nrow(filter(uni_concreteness, is.na(concreteness))) - 2` CDI words that weren't part of the collected norms (mostly animal sounds such as _baa baa_), we imputed a concreteness rating from the mean of all CDI words' concreteness rating.

### Valence and Arousal

We also used previously collected norms for valence and arousal [@warriner2013], for which adult participants are asked to rate words on a 1-9 happy-unhappy scale (valence) and 1-9 excited-calm scale (arousal). For the `r nrow(filter(uni_valences, is.na(valence))) - 2` CDI words that weren't part of the collected norms (mostly function words such as _her_), we imputed ratings from the mean of all CDI words' ratings.

### Babiness

Lastly, we used previously collected norms of "babiness", a measure of association with infancy [@perry2015] for which adult participants are asked to judge how relevant to babies a word is.

# Analysis

An overview of our entire dataset can be seen in Figure \ref{fig:data}, which shows each word's estimated age of acquisition against its predictor values, separated by language and lexical category. We present three analyses of these data: 1) how predictor values change for words learned earlier and later, 2) their relative contributions to predicting AoA, and 3) their interaction with lexical category.

## Developmental Trajectory

To assess developmental trends, we examine how the values of each predictor change as a function of estimated AoA. Figure \ref{fig:devo} shows these trajectories, with a cubic curve smoothing over all words. Words that are learned earlier are more frequent, higher in babiness, and appear in shorter sentences. Concreteness exhibits a U-shaped trajectory, with the earliest learned words actually being relatively abstract, such as many social routines and animal sounds.

## Predicting AoA

We fit a linear regression for each language's data, as well as a linear mixed-effects model with language as a random effect for all the data pooled. For illustrative purposes, Figure \ref{fig:fit} compares the predictions of the model to AoA estimates, for only English data, with outliers labelled.

Figure \ref{fig:coefs} shows the magnitude and direction of the coefficient for each predictor in each language and cross-linguistically. We find that frequency, babiness, concreteness, and MLU are relatively stronger predictors of age of acquisition, across languages and in the cross-linguistic model. Overall there's considerable consistency in how the predictors pattern in various languages, although with interesting differences: for example, MLU in English appears to be unusually strong, while frequency in Spanish look unusually weak. There is also variability in the overall fit of the models to the data, with some languages, such as Norwegian, having relatively more of the variance explained than others, such as Turkish.

## Lexical Category

Previous work gives reason to believe that predictors' relationship with age of acquisition differs among various lexical categories [@goodman2008]. To investigate these effects, we break down our data by lexical category and fit separate cross-linguistic linear mixed-effects models for each one. Figure \ref{fig:coefs_lexcat} shows the magnitudes and directions of the resulting coefficients, leaving off the less strong predictors. We find that frequency matters most for Nouns and comparatively little for Function Words, while MLU is irrelevant for both Nouns and Predicates, but highly informative for Function Words and Other items.

_[TODO: mention predictors' correlations to each other]_

_[TODO: mention that predictors have different variabilities, that probably matters]_

_[TODO: emphasize the sketchiness of using ratings cross-linguistically]_

# Discussion

What makes words easier or harder for young children to learn? Previous experimental work has largely addressed this question using small-scale experiments. While such experiments can identify sources of variation, they typically do not allow for different sources to be compared in detail. In contrast, observational studies allow the effects of individual factors (with frequency being the most common) to be measured across ages and lexical categories [e.g., @goodman2008]. Scale comes at a cost in terms of detail, however: Both the predictors and outcome data available have been quite limited. 

By including `r length(unique(lang_coef$language))` languages and as many predictors, our current work expands the scope of previous observational studies of age of acquisition. Our data show a number of patterns that confirm and expand previous reports. First, predictors differed in importance across even very early development. For example, certain concepts that were more strongly associated with babies appeared to be learned early for children across languages [@tardif2008]. Second, we found general consistency in predictor coefficients across languages (even as overall model fit varied, at least in part due to the amount and quality of data for different languages). This consistency supports the idea that differences in culture or language structure do not lead to fundamentally different acquisition strategies. 

Finally, different predictors appeared more important across lexical categories. Frequent, concrete nouns were learned earlier, consistent with theories that emphasize the importance of early referential speech [e.g., @baldwin1995]. But for predicates, concreteness was somewhat less important, and for function words, MLU was most predictive. Overall thse findings are consistent with theories that emphasize the role of linguistic structure over conceptual complexity in the acquisition of other lexical categories beyond nouns [@gentner2001, @snedeker2007]. 

Despite its larger scope, our work still shares a number of important limitations with previous studies. First and foremost, our approach is to predict one set of individuals with data about the experience of a completely different set and ratings of concepts gathered from yet others. In contrast to dense-data approaches [@roy2015], this approach fundamentally limits the amount of variability we will be able to capture. In addition, the granularity of the predictors that can be extracted from corpus data and applied to every word is necessarily quite coarse. Ideally, predictors could be targeted more specifically at particular theoretical constructs of interest (for example, the patterns of use for particular predicates).

Perhaps the most important theoretical challenge in the study of early language is how to connect between the scale of an individual observation or experiment and the broader patterns of acquisition that we observe in large datasets. We have strong theories of how individual learning situations proceed [@frank2009, @mcmurray2012]. We may not yet be able to discern the unambiguous signatures of these theories in the aggregate data we have available. But we believe that searching for these signatures at scale is a critical step in making progress on understanding language learning.  

# Acknowledgements

Thanks to the MacArthur-Bates CDI Advisory Board. This work supported by NSF BCS #1528526.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
